---
title: "Final Project RMD"
author: "Francesca Chiappetta, Malkie Wall, Chrissie Pantoja"
date: "12/2/2021"
output:
  html_document: default
  pdf_document: default
---

**Response variables**
Change in percent forest area (% of land area) - constructed from https://data.worldbank.org/indicator 

**Explanatory variables**
Forest area (sq km) - https://data.worldbank.org/indicator 
Agricultural land (% of land area) - https://data.worldbank.org/indicator
Population density - https://data.worldbank.org/indicator 
GDP per capita  - https://data.worldbank.org/indicator 

Direct democracy (mean years school, impartial adminsitration attributes) - https://govdata360.worldbank.org/indicators/hc7b6c41f?country=BRA&indicator=41865&viz=line_chart&years=1975,2019

### Read in Data
```{r libraries, include=FALSE}
library(wbstats) #package to get worldbank data
library(rio)
library(tidyverse)
library(tidyr)
library(stringr) #used to clean up strings of text.
library(openxlsx)
library(panelr)
library(dplyr)
library(knitr)
library(reshape) # a package to melt data
library(moments) #find skewness
library(gt) # a package to make tables
library(psych) #package to make panel plots 
library(GGally) #scatterplot matrix
library(plotly)
library(ggplot2)
library(skimr)
library(plm)
library(sandwich) # a package that includes functions for robust standard errors 
library(car) # a companion to applied regression (car), functions to test assumptions
library(olsrr) #Cook's Distance test
library(broom) #allows us to get model fit statistics
library(lmtest)
library(clubSandwich)
library(DescTools)
library(ltm)
library(huxtable) #package to create LaTeX and HTML tables
library("reshape2") #melt function
library(gtsummary)

set.seed(42)
# define the markup language we are working in.
# options(qwraps2_markup = "latex") is also supported.
options(qwraps2_markup = "markdown")

#remove scientific notation
options(scipen=999)

```

```{r read data, results=FALSE, echo=FALSE}
#I think delete the lines that are hashed out because they don't work 
forest.percent <- wb_data(indicator = "AG.LND.FRST.ZS")  #AG.LND.FRST.ZS 
forest.area <- wb_data(indicator = "AG.LND.FRST.K2") #AG.LND.FRST.K2
ag.percent <- wb_data(indicator = "AG.LND.AGRI.ZS") #ag.percent.code
pop.density <- wb_data(indicator = "EN.POP.DNST") #EN.POP.DNST
GDPpc <- wb_data(indicator = "NY.GDP.PCAP.CD") #NY.GDP.PCAP.CD
forest.depletion <- wb_data(indicator = "NY.ADJ.DFOR.GN.ZS")#NY.ADJ.DFOR.GN.ZS

direct.democracy <- read.csv("https://govdata360-backend.worldbank.org/api/v1/datasets/4034/dump.csv")
```

```{r data extraction, results=FALSE, echo=FALSE}
head(direct.democracy)
#pull out data from direct democracy csv



#41944 - impartial administration
impartial.administration <-direct.democracy %>% 
    filter(Indicator.Id==41944)

#41794 - absence of corruption 
absence.corruption<-direct.democracy %>% 
    filter(Indicator.Id==41794)

#42031 - predictable enforcement
predictable.enforcement<-direct.democracy %>% 
    filter(Indicator.Id==42031)

#41983 - mean years of schooling
years.schooling <-direct.democracy %>% 
    filter(Indicator.Id==41983)


```

### Joining and cleaning data 
```{r create forest.percent.change variable, results=FALSE, echo=FALSE}

#create forest percent change variable 
head(forest.percent)

forest.percent.wide <- spread(forest.percent, date, AG.LND.FRST.ZS)

forest.percent.wide$forest.percent.change.2010 <- forest.percent.wide$`2010` - forest.percent.wide$`2009`
forest.percent.wide$forest.percent.change.2011 <- forest.percent.wide$`2011` - forest.percent.wide$`2010`
forest.percent.wide$forest.percent.change.2012 <- forest.percent.wide$`2012` - forest.percent.wide$`2011`
forest.percent.wide$forest.percent.change.2013 <- forest.percent.wide$`2013` - forest.percent.wide$`2012`
forest.percent.wide$forest.percent.change.2014 <- forest.percent.wide$`2014` - forest.percent.wide$`2013`
forest.percent.wide$forest.percent.change.2015 <- forest.percent.wide$`2015` - forest.percent.wide$`2014`
forest.percent.wide$forest.percent.change.2016 <- forest.percent.wide$`2016` - forest.percent.wide$`2015`
forest.percent.wide$forest.percent.change.2017 <- forest.percent.wide$`2017` - forest.percent.wide$`2016`
forest.percent.wide$forest.percent.change.2018 <- forest.percent.wide$`2018` - forest.percent.wide$`2017`
forest.percent.wide$forest.percent.change.2019 <- forest.percent.wide$`2019` - forest.percent.wide$`2018`

head(forest.percent.wide)
glimpse(forest.percent.wide)
```

```{r pivot forest percent change data, results=FALSE, echo=FALSE}
#forest percent change
forest.percent.change.long <-forest.percent.wide %>%
  dplyr::select(3, 69:77) %>%
         pivot_longer(!country, names_to = "date", values_to = "forest.percent.change")
```

```{r look at forest percent change data, results=FALSE, echo=FALSE}
head(forest.percent.change.long)
```

```{r edit forest percent change data, results=FALSE, echo=FALSE}
forest.percent.change.long$date<-as.numeric(str_remove(forest.percent.change.long$date, "forest.percent.change."))
head(forest.percent.change.long)
```

```{r pivot democracy & schooling data, results=FALSE, echo=FALSE}
#get rid of unnecessary data and pivot long and fix year

#impartial administration (attribute)
impartial.administration.long <-impartial.administration %>%
  dplyr::select(!c(Country.ISO3, Indicator.Id, Indicator,Subindicator.Type)) %>%
         pivot_longer(!Country.Name, names_to = "date", values_to = "impartial.administration")

impartial.administration.long$date<-as.numeric(str_remove(impartial.administration.long$date, "X"))
names(impartial.administration.long)[1] <- "country"

#absence of corruption (subattribute)
absence.corruption.long <-absence.corruption %>%
  dplyr::select(!c(Country.ISO3, Indicator.Id, Indicator,Subindicator.Type)) %>%
         pivot_longer(!Country.Name, names_to = "date", values_to = "absence.corruption")

absence.corruption.long$date<-as.numeric(str_remove(absence.corruption.long$date, "X"))
names(absence.corruption.long)[1] <- "country"

#predictable enforcement (subattribute)
predictable.enforcement.long <- predictable.enforcement %>%
  dplyr::select(!c(Country.ISO3, Indicator.Id, Indicator,Subindicator.Type)) %>%
         pivot_longer(!Country.Name, names_to = "date", values_to = "predictable.enforcement")

predictable.enforcement.long$date<-as.numeric(str_remove(predictable.enforcement.long$date, "X"))
names(predictable.enforcement.long)[1] <- "country"

#mean years of schooling 
years.schooling.long <-years.schooling %>%
  dplyr::select(!c(Country.ISO3, Indicator.Id, Indicator,Subindicator.Type)) %>%
         pivot_longer(!Country.Name, names_to = "date", values_to = "years.schooling")

years.schooling.long$date<-as.numeric(str_remove(years.schooling.long$date, "X"))
names(years.schooling.long)[1] <- "country"




```

```{r clean control variable data, results=FALSE, echo=FALSE}
forest.percent2 <- dplyr::select(forest.percent, -1, -2, -6, -7, -8, -9)
head(forest.percent2)

forest.area2 <- dplyr::select(forest.area, -1, -2, -6, -7, -8, -9)
head(forest.area2)

ag.percent2 <- dplyr::select(ag.percent, -1, -2, -6, -7, -8, -9)
head(ag.percent2)

pop.density2 <- dplyr::select(pop.density, -1, -2, -6, -7, -8, -9)
head(pop.density2)

GDPpc2 <- dplyr::select(GDPpc, -1, -2, -6, -7, -8, -9)
head(GDPpc2)

forest.depletion2 <- dplyr::select(forest.depletion, -1, -2, -6, -7, -8, -9)

```

```{r join data, results=FALSE, echo=FALSE}
#join forest.percent2 and forest.area2
master.df<-inner_join(forest.percent2, forest.area2, by=c("country"="country", "date"="date"))

#join master.df and ag.percent2
master.df<-inner_join(master.df, ag.percent2, by=c("country"="country", "date"="date"))

#join master.df and pop.density2
master.df<-inner_join(master.df, pop.density2, by=c("country"="country", "date"="date"))

#join master.df and GDPpc2
master.df<-inner_join(master.df, GDPpc2, by=c("country"="country", "date"="date"))

#join master.df and years.schooling
master.df<-inner_join(master.df, years.schooling.long, by=c("country"="country", "date"="date"))

#join master.df and forest.percent.change.long
master.df<-inner_join(master.df, forest.percent.change.long, by=c("country"="country", "date"="date"))

#join master.df and impartial.administration.long
master.df<-inner_join(master.df, impartial.administration.long, by=c("country"="country", "date"="date"))

#join master.df and predictable.enforcement.long
master.df<-inner_join(master.df, predictable.enforcement.long, by=c("country"="country", "date"="date"))

#join master.df and absence.corruption.long
master.df<-inner_join(master.df, absence.corruption.long, by=c("country"="country", "date"="date"))


#join master df and foreest depletion
master.df<-inner_join(master.df, forest.depletion2, by=c("country"="country", "date"="date"))

head(master.df)
```

```{r rename columns on joined data, results=FALSE, echo=FALSE}
#rename columns

#AG.LND.FRST.ZS = forest.percent ; AG.LND.FRST.K2 = forest.area ; AG.LND.AGRI.ZS = ag.percent ; EN.POP.DNST = popn.density ; NY.GDP.PCAP.CD = GDP.per.capita 

colnames <- c("country", "year", "forest_percent", "forest_area", "ag_percent", "pop_density", "GDP_per_capita", "years_schooling", "forest_percent_change", "impartial_admin", "predictable_enforcement", "absence_corruption", "forest_depletion")

names(master.df) <- colnames

#check dataframe
head(master.df)

```
```{r dataframe of missingness across country-year }
missing.bycountry<-master.df %>% 
  group_by(country, year) %>% 
  summarise(sumNA.forest.percent = sum(is.na(forest_percent )),
            sumNA.forest.area = sum(is.na(forest_area)),
            sumNA.ag.percent = sum(is.na(ag_percent)),
            sumNA.pop.density = sum(is.na(pop_density)),
            sumNA.GDP.per.capita = sum(is.na(GDP_per_capita)),
            sumNA.years.schooling = sum(is.na(years_schooling)),
            sumNA.forest.percent.change = sum(is.na(forest_percent_change)),
            sumNA.impartial.admin = sum(is.na(impartial_admin)),
            sumNA.predictable.enforcement = sum(is.na(predictable_enforcement)),
            sumNA.absence.corruption = sum(is.na(absence_corruption)),
            ) %>%
  filter(sumNA.forest.percent>0 | 
           sumNA.forest.area>0 |
           sumNA.ag.percent>0 |
           sumNA.pop.density>0 |
           sumNA.GDP.per.capita>0 |
         sumNA.years.schooling>0 |
            sumNA.forest.percent.change>0 |
            sumNA.impartial.admin>0 |
            sumNA.predictable.enforcement>0 |
           sumNA.absence.corruption>0
         )

missing.bycountry

```

```{r look at data, results=FALSE, echo=FALSE}
#glimpse
glimpse(master.df)

#check na
sapply(master.df, function(x) sum(is.na(x)))
```



```{r cronbach alpha, results=FALSE, echo=TRUE}
#FROM MALKIE -- Not sure what this code is. @Frankie can you check whether it still makes sense? I noticed there is a master2.df and we currently only have a master.df

cronbach.alpha(master.df[,c(11,12)], na.rm = TRUE)

#higher the value, the more the variables in the index tend to correlate. A rule of thumb is that 0.7 indicates that the scale is "reliable" c
#in case the total Cronbach’s Alpha value is below the acceptable cut-off of 0.7 (mostly if an index has few items), the mean inter-item-correlation is an alternative measure to indicate acceptability; satisfactory range lies between 0.2 and 0.4
```


### Exploratory Data Analysis

*WE ONLY NEED ONE OF THE FOLLOWING TWO CHUNKS (SEE NOTE). DELETE THE OTHER ONE.* 
```{r summary table, message=FALSE, results=TRUE, echo=FALSE}

#convert to data frame because its in a strange format because of dyplr
master.df <- as.data.frame(master.df)

master.df.long <- melt(data = master.df, id.vars = c("country", "year"))

summary.master.df.long <- master.df.long %>% 
  dplyr::group_by(variable) %>%
  dplyr::select(value) %>%
  dplyr::summarize(
                    length_value =length(value) - sum(is.na(value)) , #length of value for each variable minus the sum of NA (2)
                    mean_value=mean(value, na.rm = TRUE), #remove NA
                    median_value=median(value, na.rm = TRUE), #remove NA
                    min_value=min(value, na.rm=TRUE), #remove NA
                    max_value=max(value, na.rm = TRUE), #remove NA
                    sd_value=sd(value, na.rm = TRUE), #remove NA
                    skew_value=skewness(value, na.rm = TRUE), #remove NA
                    )

# report source
source <- "Data Source: World Bank"

# make gt summary box
summary.master.df.long %>%
gt() %>%
  tab_header(
    title = md("Summary Statistics"),
    subtitle = " ") %>% tab_source_note(
    source_note = source) %>%
  
 
  fmt_passthrough (columns=c(variable)) %>%
  fmt_number(columns = c(length_value), decimals = 2) %>%
  fmt_number(columns = c(mean_value), decimals= 2) %>%
  fmt_number(columns = c(median_value), decimals = 2) %>%
  fmt_number(columns = c(min_value), decimals = 2) %>%
  fmt_number(columns = c(max_value), decimals = 2) %>%
  fmt_number(columns = c(sd_value), decimals = 2) %>%
  fmt_number(columns = c(skew_value), decimals = 2) %>%

cols_label(
    variable="Variable",
    length_value = "Observations",
    mean_value = "Mean",
    median_value = "Median",
    min_value = "Minimum",
    max_value = "Maximum",
    sd_value = "SD",
    skew_value = "Skewness")

```

#### Scatterplots


```{r log variables, echo = FALSE}
#Based on the scatterplots above, we decided to transform a few variables
#logged pop.density, gddpc, & forest.area

#population.density
master.df$log_pop_density <- log(master.df$pop_density)

#GDP.per.capita
master.df$log_gdppc <- log(master.df$GDP_per_capita)

#forest.area
master.df$log_forest_area <- log(master.df$forest_area + .0000001)


glimpse(master.df)
```


```{r model1 correlation matrix, message=FALSE, results=FALSE, echo=FALSE, fig.height=10, fig.width=10}
# @Frankie - Add in the columns you want for our model1 (including any transformed variables). Right now, I've included logged pop.density, forest.area, and gdppc

pairs.panels(master.df[c(14:16,5,8,10:12, 9)], #response variable should be in last column
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             pch=".",
             density = TRUE,  # show density plots
             ellipses = FALSE)# show correlation ellipses


```

*NOTE FROM MALKIE: The scatterplot matrix shows a very strong correlation among our impartial administration variables, so we should probably only include one at a time in our model...*



```{r subset pos and neg deforestation}
#CREATE DEFOFRESTATION DATA FRAME
#postive deforestion (NEGATIVE FOREST AREA CHANGE)
df_positive_def_rate <- master.df[master.df$forest_percent_change<0,]

#negative deforestation (POSITIVE FOREST AREA CHANGE)
df_negative_def_rate <- master.df[master.df$forest_percent_change>0,]

```


## Models

### MODEL 1- ALL COUNTRIES USING IMPARTIAL_ADMIN USING FIXED EFFECTS
```{r FIXED EFFECTS MODEL}
FE_all <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = master.df,
                   index = c("country", "year"),
                  model = "within")
summary(FE_all)
```

#### ALL COUNTRIES USING IMPARTIAL_ADMIN USING RANDOM EFFECTS
```{r RANDOM EFFECTS model }
RE_all <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = master.df,
                   index = c("country", "year"),
                  model = "random")
summary(RE_all)
```

##### TESTING
```{r PHTEST}
phtest(FE_all, RE_all)

#Ho: RANDOM EFFECTS is consistent
#Ha FE is consistant

##If p value is < 0.05 then use fixed effects
```


```{r model1 breusch pagan test}
#Testing for homoskedasticity
plmtest(RE_all, c("time"), type=c("bp"))

#he null hypothesis that heteroskedasticity is not present
#failed to reject null
# DONT HAVE HETEROSKEDASTICITY

```


## Model 1a - JUST DEFORESTATION COUNTRIES

#### FIXED EFFECTS
```{r FIXED EFFECTS MODEL DEFORESTATION ONLY}
FE_all_deforestation <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = df_positive_def_rate,
                   index = c("country", "year"),
                  model = "within")
summary(FE_all_deforestation)
```

#### DEFORESTATION COUNTRIES USING IMPARTIAL_ADMIN USING RANDOM EFFECTS
```{r RANDOM EFFECTS model deforestation all countries }
RE_all_deforestation <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = df_positive_def_rate,
                   index = c("country", "year"),
                  model = "random")
summary(RE_all_deforestation)
```
##### TESTING
```{r PHTEST deforested countries}
phtest(FE_all_deforestation, RE_all_deforestation)

#Ho: RANDOM EFFECTS is consistent
#Ha FE is consistant

##If p value is < 0.05 then use fixed effects
## CHOSE FE_ALL going forward
```


```{r model1 deforestation breusch pagan test}
#Testing for homoskedasticity
plmtest(RE_all_deforestation, c("time"), type=c("bp"))

#he null hypothesis that heteroskedasticity is not present
#failed to reject null
# DONT HAVE HETEROSKEDASTICITY

```


## Model 2 - Filter for South America

```{r South America }

master.df.SA <-master.df %>%
  filter(country=="Argentina"|country=="Bolivia"|country=="Brazil"|country=="Chile"|country=="Ecuador"|country=="Guyana"|country=="Paraguay"|country=="Peru"|country=="Suriname"|country=="Uruguay"|country=="Venezuela")
```


```{r FIXED EFFECTS MODEL SOUTH AMERICA}
FE_SA <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = master.df.SA,
                   index = c("country", "year"),
                  model = "within")

summary(FE_SA)
```

```{r model2 breusch pagan test}
#Testing for homoskedasticity
plmtest(FE_SA, c("time"), type=c("bp"))

#he null hypothesis that heteroskedasticity is not present
#failed to reject null
# DONT HAVE HETEROSKEDASTICITY

```


### SUBSETTING TROPICAL COUNTRIES
```{r subsetting tropical countries}
library(data.table)
tropical_countries <- fread("/Users/frankiechiappetta/Desktop/csvData.csv")
tropical_countries

tropical_countries2 <- setDT(master.df)[country %chin% tropical_countries$country]
tropical_countries2
```

```{r FIXED EFFECTS MODEL TROPICAL COUNTRIES}
FE_tropical <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = tropical_countries2,
                   index = c("country", "year"),
                  model = "within")
summary(FE_tropical)
```


```{r RANDOM EFFECTS model TROPICAL COUNTRIES}
RE_tropical <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = tropical_countries2,
                   index = c("country", "year"),
                  model = "random")
summary(RE_tropical)
```

```{r PHTEST tropical countries }
phtest(FE_tropical, RE_tropical)

#Ho: RANDOM EFFECTS is consistent
#Ha FE is consistant

## CHOSE RE going forward
```


```{r breusch pagan test tropical countries }
#Testing for homoskedasticity
plmtest(RE_tropical, c("time"), type=c("bp"))

#he null hypothesis that heteroskedasticity is not present
#failed to reject null
# DONT HAVE HETEROSKEDASTICITY

```



```{r DEFORESTED TROPICAL COUNTRIES}

tropical_countries_deforested <- setDT(df_positive_def_rate)[country %chin% tropical_countries$country]
tropical_countries_deforested

```


```{r FIXED EFFECTS MODEL DEFORESTED TROPICAL COUNTRIES}
FE_tropical_deforested <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = tropical_countries_deforested,
                   index = c("country", "year"),
                  model = "within")
summary(FE_tropical_deforested)
```


```{r RANDOM EFFECTS model DEFORESTED TROPICAL COUNTRIES}
RE_tropical_deforested <- plm(forest_percent_change ~ impartial_admin + years_schooling + ag_percent + log_forest_area + log_gdppc + I(log_gdppc^2) + log_pop_density,
                   data = tropical_countries_deforested,
                   index = c("country", "year"),
                  model = "random")
summary(RE_tropical_deforested)
```

```{r PHTEST deforested tropical countries }
phtest(FE_tropical_deforested, RE_tropical_deforested)

#Ho: RANDOM EFFECTS is consistent
#Ha FE is consistant

## CHOSE RE going forward becuase p > 0.05
```


```{r breusch pagan test deforested tropical countries }
#Testing for homoskedasticity
plmtest(RE_tropical_deforested, c("time"), type=c("bp"))

#he null hypothesis that heteroskedasticity is not present
#failed to reject null
# DONT HAVE HETEROSKEDASTICITY

```
